Between Subgraph Isomorphism and Maximum Common Subgraph

### 2016-08-26 paper/paper.tex
Showframe hackery, for now
New experiments, and use gpgnode results
General format tinkering
Some runtimes
Histogrammification
Plotting
Words words words

The subgraph isomorphism problem is to find a copy of a small \emph{pattern} graph inside a larger
\emph{target} graph. It comes in two forms: in the non-induced variant, edges must be mapped to
edges, but the target may have ``extra edges'', whilst in the induced variant, non-edges may only be
mapped to non-edges. When a pattern cannot be found, we may wish to be given a result which maps as
many vertices of the pattern into the target as possible. In the induced case, this is known as the
maximum common induced subgraph problem (we discuss the non-induced case below). However, although
recent subgraph isomorphism algorithms are comfortable working with graphs with thousands of
vertices, the state of the art for the maximum common subgraph problem becomes computationally
infeasible at only 35 vertices when working with unlabelled graphs \citep{CP2016MCSPaper}. This is
largely because strong inference, based upon the degrees of vertices
\citep{DBLP:journals/ai/Solnon10} and the distances or paths between them
\citep{DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15}, is possible with subgraph isomorphism,
but not maximum common subgraph, and so the state space in the former is much more restricted. In
this work we discuss an intermediate problem, where we must map all but $k$ vertices of the pattern
graph into the target. We show that if $k$ is reasonably small (say, between 1 and 5), then weakened
forms of degree and path based filterings are still effective in pruning the search space and
providing additional constraints respectively. We then show that combining these techniques does
lead to a practical algorithm which can scale to work with the families of graphs commonly used to
benchmark subgraph isomorphism algorithms.

@@

A non-induced subgraph isomorphism from a graph $P$ to a graph $T$ is an injective mapping
$P \rightarrowtail T $ which maps adjacent vertices to adjacent vertices. An induced subgraph
isomorphism $P \hookrightarrow T$, additionally maps non-adjacent vertices to non-adjacent
 vertices.

A common induced subgraph of graphs $G$ and $H$ is a graph $P$, together with two induced subgraph
isomorphisms to $G$ and $H$; a maximum common induced subgraph is a common induced subgraph with as
many vertices as possible.

A maximum common partial (i.e.\ non-induced) subgraph would allow us to select every vertex in the
pattern, and none of the edges. The objective is therefore changed to maximise the number of edges,
rather than vertices. However, this is not what we are looking for: usually maximum common subgraph
problems are symmetric in their inputs, but we are allowing extra edges only in the target, not in
the pattern.

We define a $k$-less-subgraph isomorphism from $P$ to $T$ to be a subgraph isomorphism from all
but $k$ vertices of $P$ to $T$; this may be non-induced or induced, written
$P \lessnonind{k} T$ and $P \lessind{k} T$ respectively.

\subsection{Models and Algorithms}

@@

Comparison with the clique and CP models, as well as ``drop k vertices and try each in turn''.

@@

@@

Just results and ``drop k vertices and try each in turn''.

@@

\subsection{Parallel Search}

>>>>

The subgraph isomorphism problem is to find a copy of a small \emph{pattern} graph inside a larger
\emph{target} graph. It comes in two forms: in the non-induced variant, edges must be mapped to
edges, but the target may have ``extra edges'', whilst in the induced variant, non-edges may only be
mapped to non-edges. When a pattern cannot be found, we may wish to be given a result which maps as
many vertices of the pattern into the target as possible. In the induced case, this is known as the
maximum common induced subgraph problem (we discuss the non-induced case below). However, although
recent subgraph isomorphism algorithms are comfortable working with graphs with thousands of
vertices, the state of the art for the maximum common subgraph problem becomes computationally
infeasible at only 35 vertices when working with unlabelled graphs \citep{CP2016MCSPaper}. This is
largely because strong inference, based upon the degrees of vertices
\citep{DBLP:journals/ai/Solnon10} and the distances or paths between them
\citep{DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15}, is possible with subgraph isomorphism,
but not maximum common subgraph, and so the state space in the former is much more restricted. In
this work we discuss an intermediate problem, where we must map all but $k$ vertices of the pattern
graph into the target. We show that if $k$ is reasonably small (say, between 1 and 5), then weakened
forms of degree and path based filterings are still effective in pruning the search space and
providing additional constraints respectively. We then show that combining these techniques does
lead to a practical algorithm which can scale to work with the families of graphs commonly used to
benchmark subgraph isomorphism algorithms: depending upon the benchmark family, in a substantial
portion of ; in many more cases, we can at least obtain an upper bound.

@@

A non-induced subgraph isomorphism from a graph $P$ to a graph $T$ is an injective mapping $P
\rightarrowtail T $ which maps adjacent vertices to adjacent vertices. An induced subgraph
isomorphism $P \hookrightarrow T$ additionally maps non-adjacent vertices to non-adjacent vertices.
We define a $k$-less-subgraph isomorphism from $P$ to $T$ to be a subgraph isomorphism from all but
$k$ vertices of $P$ to $T$; this may be non-induced or induced, written $P \lessnonind{k} T$ and $P
\lessind{k} T$ respectively.

A common induced subgraph of graphs $G$ and $H$ is a graph $P$, together with two induced subgraph
isomorphisms to $G$ and $H$; a maximum common induced subgraph is a common induced subgraph with as
many vertices as possible. An induced $k$-less subgraph isomorphism $P \lessind{k} T$ is equivalent
to a common induced subgraph of $P$ and $T$ with $|\operatorname{V}(P)| - k$ vertices.

If defined similarly, a maximum common non-induced subgraph would allow us to select every vertex in
the smaller of the two graphs, and none of the edges. It is therefore traditional to change the
objective to maximise the number of edges selected, rather than vertices, when a non-induced common
subgraph is sought---this problem is usually called the maximum common \emph{partial} subgraph
problem instead. However, this is not what we will be discussing in this paper: maximum common
subgraph problems are symmetric in their inputs, but when discussing the non-induced case we are
allowing extra edges only in the target graph, not in the pattern.

?? Draw an example.

\subsection{Constraint Models and Algorithms}

@@

@@

\begin{figure}
    \includegraphics*{gen-graph-which-k-by-family.pdf}
    \caption{There are satisfiable instances for small values of $k$. Yay.}\label{figure:which-k}
\end{figure}

@@

@@

\subsection{Solving from the Top Down}

\begin{figure}
    \includegraphics*{gen-graph-versus-cp.pdf}
    \caption{A righteous spanking.}\label{figure:versus-cp}
\end{figure}
